Decision entropy

### 2023-01-18 analysis/decisionEntropy.Rmd
click entropy analysis

# What  I refer to as 'prior' here is the overall tendency of participants to
# click on some cells. To make sure this quantity is not driven by the content
# of the board itself, for each board I compute this quantity across all other
# boards, not including itself.
E2.prior_from_group <- E2.click_log_with_boards %>% 
  filter(test_part=='nonpretend') %>% # we are interested in true games
  mutate(flat_position = factor(flat_position, levels=seq(25))) %>%
  group_by(grid_number,click_number, flat_position) %>%
  summarise(n=n())%>% # count the number of clicks per cell per board per timepoint
  group_by(click_number, flat_position)%>%
  ## for each grid, average across all other grids.
  do( data.frame(.,other_n=sapply(.$grid_number, function(i) mean(.$n[!.$grid_number %in% i])))) %>%
  group_by(grid_number, click_number) %>%
  mutate(p=other_n/sum(other_n)) %>% # normalize to get a probability function
  dplyr::select(grid_number,click_number,flat_position,p) %>%
  spread(flat_position,p, fill=0, sep='_') %>% # pivot to wider format
  ungroup();

@@

>>>>

# What  I refer to as 'prior' here is the overall tendency of participants to
# click on some cells. To make sure this quantity is not driven by the content
# of the board itself, for each board I compute this quantity across all other
# boards, not including itself.
E2.prior_from_group <- E2.click_log_with_boards %>% 
  filter(test_part=='nonpretend') %>% # we are interested in true games
  mutate(flat_position = factor(flat_position, levels=seq(25))) %>%
  group_by(grid_number,click_number, flat_position) %>%
  summarise(n=n())%>% # count the number of clicks per cell per board per timepoint
  group_by(click_number, flat_position)%>%
  ## for each grid, sum across all other grids.
  do( data.frame(.,other_n=sapply(.$grid_number, function(i) sum(.$n[!.$grid_number %in% i])))) %>%
  group_by(grid_number, click_number) %>%
  mutate(p=other_n/sum(other_n)) %>% # normalize to get a probability function
  dplyr::select(grid_number,click_number,flat_position,p) %>%
  spread(flat_position,p, fill=0, sep='_') %>% # pivot to wider format
  ungroup();

@@

```{r entropy-by-erial-position, echo=FALSE, cache=TRUE}

#compute Shannon entropy
entropy <- function(target) {
  freq <- table(target)/length(target)
  # vectorize
  vec <- as.data.frame(freq)[,2]
  #drop 0 to avoid NaN resulting from log2
  vec<-vec[vec>0]
  #compute entropy
  -sum(vec * log2(vec))
}

E2.click_entropy_by_serial_position <- E2.click_log_with_boards %>%
  rbind(E2.click_log_with_boards_optimal)%>%
  rbind(E2.click_log_with_boards_random)%>%
  # filter(click_number<15) %>%
  group_by(subj_id,test_part,click_number)%>%
  summarise(entropy=entropy(flat_position))

E2.click_entropy_by_serial_position_summary <- E2.click_entropy_by_serial_position %>%
  group_by(test_part,click_number) %>%
  summarise(mean_entropy=mean(entropy),
            se_entropy=se(entropy)) %>%
  mutate(test_part = factor(test_part, levels=c('nonpretend','pretend','random','optimal')));

p <- E2.click_entropy_by_serial_position_summary %>%
  ggplot(aes(x=click_number,y=mean_entropy,color=test_part,group=test_part)) +
  geom_line() +
  geom_ribbon(aes(ymax=mean_entropy+se_entropy,ymin=mean_entropy-se_entropy, fill=test_part),alpha=0.5) +
  scale_fill_manual(values=c("#69b3a2", "#404080", "#FDE725", "black")) +
  scale_color_manual(values=c("#69b3a2", "#404080", "black", "black")) +
  labs(x='click number', y='entropy') +
  theme_classic()

ggsave('../figures/E2_entropy_by_click_nnumber.png',p,width=5,height=5,dpi=300)
